### 2022-09-09 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_julia_for_ecologists@42afb09366d957a043e5b21e181d703a6ebdeaf7 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix2}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\pagestyle{plain}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Julia for Ecologists / 10 Rules for
learning/using Julia as an Ecologist}
\vskip 5em

%
\href{https://orcid.org/0000-0002-6506-6487}{Michael D.\,Catchen}%
%
\,\textsuperscript{1,2}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{3,2}

\textsuperscript{1}\,McGill University\quad \textsuperscript{2}\,Québec
Centre for Biodiversity Sciences\quad \textsuperscript{3}\,Université de
Montréal

\textbf{Correspondance to:}\\
Michael D. Catchen --- \texttt{michael.catchen@mcgill.ca}\\

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill
Julia is a good language, ecologists should learn it.

\vfill

\clearpage
\linenumbers
\pagestyle{normal}

\hypertarget{outline}{%
\subsection{Outline}\label{outline}}

\begin{itemize}
\tightlist
\item
  Why should ecologists learn julia?

  \begin{itemize}
  \tightlist
  \item
    Well, there are the criteria that are directly measureble that make
    it better than R/Python:

    \begin{itemize}
    \tightlist
    \item
      fast
    \item
      native support on GPUs
    \end{itemize}
  \item
    But there are also criteria that are more subjective, and that take
    experience and practice using the language to appreciate

    \begin{itemize}
    \tightlist
    \item
      clever use of dispatch patterns
    \item
      use of one-lienrs
    \item
      using parameterized types well
    \end{itemize}
  \item
    You will learn how to be a better programmer in \emph{any} language,
    because smart use of julia requires understanding some fundemantal
    concepts in programming that are `hidden' from users in R/python
  \item
    The biggest reason \emph{not} to use julia is that the
    ecology/evolution package ecosystem in R is larger, and the ML
    ecosystem in python is more popular. However:

    \begin{itemize}
    \tightlist
    \item
      you can call \emph{any} R/python function/library using
      RCall/PyCall in julia
    \item
      More packages isn't necessarily better when they don't work
      together
    \end{itemize}
  \end{itemize}
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

10 rules to use julia effectively (highly tenative)

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Use concrete types
\item
  Use abstract types
\item
  Learn about dispatch
\item
  Understand memory allocation
\item
  Use broadcasting
\item
  Learn about the statistics ecosystem: StatsBase, Statistics, GLM, MLJ,
  Flux, Turing
\item
  Learn about the simulation libraries (DiffEq, DynamicGrids)
\item
  Learn how various statistics/simulation libraries work together
\item
\item
  Contribute to open-source community
\end{enumerate}

\end{document}

### 2022-09-09 preprint.tex
Deploying to gh-pages from @ gottacatchenall/ms_julia_for_ecologists@42afb09366d957a043e5b21e181d703a6ebdeaf7 🚀

--- 2022-09-09 draft.tex
@@
 %!TEX TS-program = xelatex
-\documentclass[11pt]{article}
+\documentclass[10pt,oneside]{article}
+\usepackage[fontsize=9pt]{scrextend}
 
@@
 \newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\usepackage[table,dvipsnames]{xcolor}
 
-\usepackage{lineno}
-\usepackage[nolists,noheads]{endfloat}
+\geometry{includemp,
+            letterpaper,
+            top=2.4cm,
+            bottom=2.4cm,
+            left=1.0cm,
+            right=1.0cm,
+            marginparwidth=5cm,
+            marginparsep=1.0cm}
+
+\usepackage[singlelinecheck=off]{caption}
+
+\captionsetup{
+  font={small},
+  labelfont={bf},
+  format=plain,
+  labelsep=quad
+}
+
+\usepackage{floatrow}
+
+\floatsetup[figure]{margins=hangright,
+              facing=no,
+              capposition=beside,
+              capbesideposition={center,outside},
+              floatwidth=\textwidth}
+
+% \floatsetup[table]{margins=hangright,
+%              facing=no,
+%              capposition=beside,
+%              capbesideposition={center,outside},
+%              floatwidth=\textwidth}
 
@@
 
-\tolerance=1
-\emergencystretch=\maxdimen
-\hyphenpenalty=10000
-\hbadness=10000
+\setcounter{secnumdepth}{5}
 
-\doublespacing
+\usepackage{titlesec}
+
+\titleformat{\section}[block]
+{\normalfont\large\sffamily}
+{\thesection}{.5em}{\titlerule\\[.8ex]\bfseries}
+
+\titleformat{\subsection}[runin]
+{\normalfont\fontseries{b}\selectfont\filright\sffamily}
+{\thesubsection.}{.5em}{}
+
+\titleformat{\subsubsection}[runin]
+{\normalfont\itshape\rmfamily\bfseries}{\thesubsubsection}{1em}{}
+
+\fancypagestyle{firstpage}
+{
+   \fancyhf{}
+   \renewcommand{\headrulewidth}{0pt}
+   \fancyfoot[R]{\footnotesize\ccby}
+   \fancyfoot[L]{\footnotesize\sffamily\today}
+}
 
@@
 }
+
+\usepackage{tikz}
 \begin{document}
-\raggedright
-\thispagestyle{empty}
-{\Large\bfseries\sffamily Julia for Ecologists / 10 Rules for
-learning/using Julia as an Ecologist}
-\vskip 5em
+\tikz [remember picture, overlay] %
+\node [shift={(-0.6in,1.1cm)},scale=0.2,opacity=0.4] at (current page.south east)[anchor=south east]{\includegraphics{logo}};%
+\pagestyle{normal}
+\thispagestyle{firstpage}
 
+\newcommand{\colorRule}[3][black]{\textcolor[HTML]{#1}{\rule{#2}{#3}}}
+
+\noindent {\LARGE \textbf{\textsf{Julia for Ecologists / 10 Rules for
+learning/using Julia as an Ecologist}}}
+
+\medskip
+\begin{flushleft}
+{\small
 %
@@
 %
-\,\textsuperscript{1,2}\quad %
+\,\textsuperscript{1,2}, %
 \href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
@@
 \,\textsuperscript{3,2}
-
-\textsuperscript{1}\,McGill University\quad \textsuperscript{2}\,Québec
-Centre for Biodiversity Sciences\quad \textsuperscript{3}\,Université de
-Montréal
-
+\vskip 1em
+\textsuperscript{1}\,McGill University; \textsuperscript{2}\,Québec
+Centre for Biodiversity Sciences; \textsuperscript{3}\,Université de
+Montréal\\
+\vskip 1em
 \textbf{Correspondance to:}\\
 Michael D. Catchen --- \texttt{michael.catchen@mcgill.ca}\\
+}
+\end{flushleft}
 
-\vfill
-This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
-Last revision: \emph{\today}
+\vskip 2em
+\makebox[0pt][l]{\colorRule[CCCCCC]{2.0\textwidth}{0.5pt}}
+\vskip 2em
+\noindent
 
-\clearpage
-\thispagestyle{empty}
+\marginpar{\vskip 1em\flushright
+{\small{\bfseries Keywords}:\par
+computational ecology\\scientific computing\\machine
+learning\\ecological statistics\\}
+}
 
-\vfill
 Julia is a good language, ecologists should learn it.
 
-\vfill
-
-\clearpage
-\linenumbers
-\pagestyle{normal}
+\vskip 2em
+\makebox[0pt][l]{\colorRule[CCCCCC]{2.0\textwidth}{0.5pt}}
+\vskip 2em
 

### 2022-09-09 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_julia_for_ecologists@9fd97f8feb8029ab7d886bcd046993b5bbeab2b4 🚀
Deploying to gh-pages from @ gottacatchenall/ms_julia_for_ecologists@10ae2c32581e8a73942f88c7a99e1118bbd15b26 🚀

@@
 \thispagestyle{empty}
-{\Large\bfseries\sffamily Julia for Ecologists / 10 Rules for
-learning/using Julia as an Ecologist}
+{\Large\bfseries\sffamily Solving the \(n\)-language problem in ecology}
 \vskip 5em
@@
 \hypertarget{outline}{%
-\subsection{Outline}\label{outline}}
+\section{Outline}\label{outline}}
 
@@
 
-10 rules to use julia effectively (highly tenative)
+\hypertarget{abstract}{%
+\section{Abstract}\label{abstract}}
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+In order to measure, understand, and mitigate the consequences of
+anthropogenic change on ecosystems the serives they provide, ecologists
+need a set of computational tools (Urban \emph{et al.} 2022). These
+tools must be performant, but crucially modular and interfaceable
+(McIntire \emph{et al.} 2022).
+
+Ecological data is often difficult to access and reuse (Gonzalez \&
+Peres-Neto 2015; Poisot \emph{et al.} 2019). Many sources of ecological,
+evolutionary, and environmental data exist, but synthesizing this data
+into a single product suitable for analysis often remains tedious as
+data are not in formats that can be easily combined or interfaced. Here
+we propose that we can solve this problem through standardization
+(Zimmerman 2008)---developing a common definition such that data
+collected in a variety of contexts can be assimilated while minimizing
+the overhead of data cleaning and wrangling.
+
+A common representation of ecological data will have three primary
+benefits: it will \textbf{1}) enable new forms of analysis by making it
+easier to combine data from different sources (Heberling \emph{et al.}
+2021), \textbf{2)} enable continuous integration of new data for
+next-generation biodiversity monitoring (Kühl \emph{et al.} 2020), and
+\textbf{3)} aid in open sharing and reproducability of published results
+(Zimmerman 2008; Borregaard \& Hart 2016). Here, we briefly review
+approaches to data standardization developed in other fields, in order
+to determine what makes an open standard succeed in promoting data
+sharing, and what doesn't. Based on the properties of good standards we
+identify, we propose building a living standard for ecological data in
+the \texttt{Julia} programming language, and argue this is necessary to
+obtain the three primary benefits of standardization mentioned earlier.
+
+The so-called ``two-language problem'' in computational science, where
+it is easier for a researcher to developer a prototype of a
+model/simulation in a high-level language, like Python or R, and later
+have to port the model to a lower-level compiled language because the
+performance of these compiled languages (e.g.~C++/Fortran) is orders of
+magnitude faster than high level interpreted languages. In fact, many of
+the most popular tools in higher-level languages are actually thin
+wrappers around a compiled (often C++) base (e.g. tidyverse, keras,
+numpy, TensorFlow, scikit-learn, pandas, etc.). However, the skills
+required to use or debug---let alone write---scientific software in
+these lower level langauges is not often taught.
+
+We propose that that Julia has certain properties absent in other
+popular languages for scientific computing that make it particularly
+suited for the development of a cohesive, modular, and extendible set of
+tools ideal for the development of a platform for ecological analysis
+{[}@{]}.
+
+\hypertarget{the-nature-of-computation-in-julia}{%
+\section{The nature of computation in
+Julia}\label{the-nature-of-computation-in-julia}}
+
+\hypertarget{types}{%
+\subsection{Types}\label{types}}
+
+Why is this useful for ecologists? Often times in ecology, the same
+information is represented in different formats. Two packages in R might
+not agree on what the ``correct'' format to represent information is.
+
+\textbf{Concept fig here}
+
+At the core of the Julia language is its \emph{type system}. Type
+systems can often be alienating to those who learned programming in
+so-called \emph{dynamically} typed languages (like R, python, and
+JavaScript). In dynamically-typed languages, \texttt{x=\ 5}, and
+\texttt{x\ =\ "hello\ world"} and the language won't care that you
+changed the type of information that was stored in \texttt{x} from a
+number to a string. Practically, this form of dynamic-typing was adopted
+because it is far more convenient to write code like that above than
+defining variables with explicit types, e.g.~how you would in C:
+\texttt{char\ c\ =\ "a";} and \texttt{int\ x\ =\ 5;}.
+
+Julia doesn't require explicit type declarations, meaning
+\texttt{x\ =\ 5} is perfectly valid code, but internallyJulia is doing
+the bookkeeping of what type of information is stored in \texttt{x},
+from an \texttt{Int64} to a \texttt{String} in the above example.
+
+Using explicit types is central to Julia's speed, but also enables much
+of its most unique and user-friendly functionality, primarily the use of
+a \emph{multiple-dispatch} system.
+
+\hypertarget{dispatch}{%
+\subsection{Dispatch}\label{dispatch}}
+
+\emph{Dispatch} refers to the way a computer program decides what
+function to call.
+
+In many staticly-typed languages, you are allow to use the same function
+name more than once.
+
+\hypertarget{doing-computational-science-in-julia}{%
+\section{\texorpdfstring{Doing computational science in
+\texttt{Julia}}{Doing computational science in Julia}}\label{doing-computational-science-in-julia}}
+
+\hypertarget{managing-data}{%
+\subsection{Managing Data}\label{managing-data}}
+
+DataFrames.jl and DFMeta.
+
+\hypertarget{doing-statistics-and-machine-learning}{%
+\subsection{Doing statistics and machine
+learning}\label{doing-statistics-and-machine-learning}}
 
@@
 \def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{6}
 \tightlist
-\item
-  Use concrete types
-\item
-  Use abstract types
-\item
-  Learn about dispatch
-\item
-  Understand memory allocation
-\item
-  Use broadcasting
 \item
@@
   Flux, Turing
+\end{enumerate}
+
+\hypertarget{doing-simulation}{%
+\subsection{Doing simulation}\label{doing-simulation}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{7}
+\tightlist
 \item
@@
 \item
-  Learn how various statistics/simulation libraries work together
-\item
-\item
-  Contribute to open-source community
+  Learn how various statistics/simulation libraries work togethe
 \end{enumerate}
 
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+Defining a living standard for ecological data in \texttt{Julia} will
+make it easier to combine data from different sources by splitting the
+process of data aggregation from the process of analysis. Integrating
+data from a particular study, or a new database, would be as simple as
+implementing the interface from the data source to the standardized
+types. Data from individual studies could be incorporated into public
+repositories containing both the raw data and the interface to Julia
+data structures, and this combined data/interface package is all that is
+needed to either reproduce the results or incorporate that particular
+study's data into analysis. This will make combining data from multiple
+sources easier, and yield benefits for the development and
+implementation of novel methods, as the software for analysis becomes
+separate from the software for data cleaning and aggregation.
+
+We envision a modern set of tools for ecology in \texttt{Julia} based
+around the standardized types. Far outside of ecology, the term
+``ecosystem'' is used metaphorically to describe a set of software tools
+that work together. We imagine multiple ``trophic-levels'' of packages
+for ecological science in \texttt{Julia} based around the ``basal'' set
+of standardized types --- a modular set of tools that can be chained
+together create arbitrarily complex analysis pipelines. that can be
+scaled to meet the needs of next-generation biodiversity monitoring.
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-Borregaard2016MorRep}{}%
+Borregaard, M.K. \& Hart, E.M. (2016). Towards a more reproducible
+ecology. \emph{Ecography}, 39, 349--353.
+
+\leavevmode\hypertarget{ref-Gonzalez2015ActSta}{}%
+Gonzalez, A. \& Peres-Neto, P.R. (2015). Act to staunch loss of research
+data. \emph{Nature}, 520, 436--436.
+
+\leavevmode\hypertarget{ref-Heberling2021DatInt}{}%
+Heberling, J.M., Miller, J.T., Noesgaard, D., Weingart, S.B. \& Schigel,
+D. (2021). Data integration enables global biodiversity synthesis.
+\emph{Proceedings of the National Academy of Sciences}, 118.
+
+\leavevmode\hypertarget{ref-Kuhl2020EffBio}{}%
+Kühl, H.S., Bowler, D.E., Bösch, L., Bruelheide, H., Dauber, J.,
+Eichenberg, David., \emph{et al.} (2020). Effective Biodiversity
+Monitoring Needs a Culture of Integration. \emph{One Earth}, 3,
+462--474.
+
+\leavevmode\hypertarget{ref-McIntire2022PerRei}{}%
+McIntire, E.J.B., Chubaty, A.M., Cumming, S.G., Andison, D., Barros, C.,
+Boisvenue, C., \emph{et al.} (2022). PERFICT: A Re-imagined foundation
+for predictive ecology. \emph{Ecology Letters}, 25, 1345--1351.
+
+\leavevmode\hypertarget{ref-Poisot2019EcoDat}{}%
+Poisot, T., Bruneau, A., Gonzalez, A., Gravel, D. \& Peres-Neto, P.
+(2019). Ecological Data Should Not Be So Hard to Find and Reuse.
+\emph{Trends in Ecology \& Evolution}, 34, 494--496.
+
+\leavevmode\hypertarget{ref-Urban2022CodLif}{}%
+Urban, M.C., Travis, J.M.J., Zurell, D., Thompson, P.L., Synes, N.W.,
+Scarpa, A., \emph{et al.} (2022). Coding for Life: Designing a Platform
+for Projecting and Protecting Global Biodiversity. \emph{BioScience},
+72, 91--104.
+
+\leavevmode\hypertarget{ref-Zimmerman2008NewKno}{}%
+Zimmerman, A.S. (2008). New Knowledge from Old Data: The Role of
+Standards in the Sharing and Reuse of Ecological Data. \emph{Science,
+Technology, \& Human Values}, 33, 631--652.
+
+\end{CSLReferences}
+
 \end{document}

### 2022-09-09 preprint.tex
Deploying to gh-pages from @ gottacatchenall/ms_julia_for_ecologists@9fd97f8feb8029ab7d886bcd046993b5bbeab2b4 🚀
Deploying to gh-pages from @ gottacatchenall/ms_julia_for_ecologists@10ae2c32581e8a73942f88c7a99e1118bbd15b26 🚀

@@
 
-\noindent {\LARGE \textbf{\textsf{Julia for Ecologists / 10 Rules for
-learning/using Julia as an Ecologist}}}
+\noindent {\LARGE \textbf{\textsf{Solving the \(n\)-language problem in
+ecology}}}
 
@@
 {\small{\bfseries Keywords}:\par
-computational ecology\\scientific computing\\machine
-learning\\ecological statistics\\}
+computational ecology\\scientific computing\\machine learning\\numerical
+ecology\\}
 }

### 2022-09-09 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_julia_for_ecologists@738048d03f64b40771406b32f43895c9d9a41177 🚀

@@
 \thispagestyle{empty}
-{\Large\bfseries\sffamily Solving the \(n\)-language problem in ecology}
+{\Large\bfseries\sffamily Solving the \(n\)-language problem: A
+ecologist's guide to learning Julia}
 \vskip 5em
@@
 
+\begin{figure}
+\centering
+\includegraphics{./figures/multiple_dispatch.png}
+\caption{TODO: caption. Adapted from Karpinski 2019 ``The unreasonable
+effectiveness of multiple dispatch''}
+\end{figure}
+
 \hypertarget{types}{%
@@
 not agree on what the ``correct'' format to represent information is.
-
-\textbf{Concept fig here}
 

### 2022-09-10 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_julia_for_ecologists@d199f6d27a498bbadc447c73c51cd36074336e9e 🚀

@@
 
-\begin{itemize}
-\tightlist
-\item
-  Why should ecologists learn julia?
+An ecologists guide to learning Julia
 
@@
 \item
-    Well, there are the criteria that are directly measureble that make
-    it better than R/Python:
+  Why should ecologists learn Julia?
 
@@
   \item
-      fast
+    No point ignoring the elephant in the room: R is by far the most
+    used language in ecology
   \item
-      native support on GPUs
-    \end{itemize}
+    But many of the problems in contemporary ecology require
+    functionality that R just can't handle (mass individual-based
+    models, big data management, quick numerical methods)
   \item
-    But there are also criteria that are more subjective, and that take
-    experience and practice using the language to appreciate
+    \texttt{julia} has particular features that solve a broad chunk of
+    the difficulties of computational analysis of ecological data
 
@@
     \item
-      clever use of dispatch patterns
-    \item
-      use of one-lienrs
-    \item
-      using parameterized types well
+      and improves user experience by enabling code that is
+      \emph{idiomatic}.
     \end{itemize}
   \item
-    You will learn how to be a better programmer in \emph{any} language,
-    because smart use of julia requires understanding some fundemantal
-    concepts in programming that are `hidden' from users in R/python
+    this makes julia well suited to become "{[}a platform for ecological
+    prediction and forecasting{]}; McIntyre etal, Urban etal
+  \end{itemize}
 \item
-    The biggest reason \emph{not} to use julia is that the
-    ecology/evolution package ecosystem in R is larger, and the ML
-    ecosystem in python is more popular. However:
+  In what order should people approach new topics
 
@@
   \item
-      you can call \emph{any} R/python function/library using
-      RCall/PyCall in julia
+    understand the core concepts that make julia different from other
+    languages first
+
+    \begin{itemize}
+    \tightlist
     \item
-      More packages isn't necessarily better when they don't work
-      together
+      Types
+    \item
+      multiple dispatch
     \end{itemize}
+  \item
+    Learn data management in Julia (emphasize analogies to tidyverse)
+
+    \begin{itemize}
+    \tightlist
+    \item
+      DataFrames
+    \end{itemize}
+  \item
+    Learn stats/ml in Julia (emphasize anologies to GLM in R and
+    scikit-learn/tensorflow in python)
+  \item
+    Learn visualization (Makie)
+  \end{itemize}
+\item
+  Discussion
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Talk about adjacent scientific computing course
+  \item
+    What does the future of computing in ecology look like?
   \end{itemize}
@@
 
-\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+\begin{longtable}[]{@{}l@{}}
+\toprule
+\endhead
+\begin{minipage}[t]{0.05\columnwidth}\raggedright
+- Why should ecologists learn julia? - Well, there are the criteria that
+are directly measurable that make it better than R/Python: - fast -
+native support on GPUs - But there are also criteria that are more
+subjective, and that take experience and practice using the language to
+appreciate - clever use of dispatch patterns - use of one-lienrs - using
+parameterized types well - You will learn how to be a better programmer
+in \emph{any} language, because smart use of julia requires
+understanding some fundemantal concepts in programming that are `hidden'
+from users in R/python - The biggest reason \emph{not} to use julia is
+that the ecology/evolution package ecosystem in R is larger, and the ML
+ecosystem in python is more popular. However: - you can call \emph{any}
+R/python function/library using RCall/PyCall in julia - More packages
+isn't necessarily better when they don't work together\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
 
@@
 In order to measure, understand, and mitigate the consequences of
-anthropogenic change on ecosystems the serives they provide, ecologists
-need a set of computational tools (Urban \emph{et al.} 2022). These
-tools must be performant, but crucially modular and interfaceable
-(McIntire \emph{et al.} 2022).
+anthropogenic change on ecosystems the services they provide, ecologists
+need a modern set of computational tools (Urban \emph{et al.} 2022).
 
-Ecological data is often difficult to access and reuse (Gonzalez \&
-Peres-Neto 2015; Poisot \emph{et al.} 2019). Many sources of ecological,
-evolutionary, and environmental data exist, but synthesizing this data
-into a single product suitable for analysis often remains tedious as
-data are not in formats that can be easily combined or interfaced. Here
-we propose that we can solve this problem through standardization
-(Zimmerman 2008)---developing a common definition such that data
-collected in a variety of contexts can be assimilated while minimizing
-the overhead of data cleaning and wrangling.
-
-A common representation of ecological data will have three primary
-benefits: it will \textbf{1}) enable new forms of analysis by making it
-easier to combine data from different sources (Heberling \emph{et al.}
-2021), \textbf{2)} enable continuous integration of new data for
-next-generation biodiversity monitoring (Kühl \emph{et al.} 2020), and
-\textbf{3)} aid in open sharing and reproducability of published results
-(Zimmerman 2008; Borregaard \& Hart 2016). Here, we briefly review
-approaches to data standardization developed in other fields, in order
-to determine what makes an open standard succeed in promoting data
-sharing, and what doesn't. Based on the properties of good standards we
-identify, we propose building a living standard for ecological data in
-the \texttt{Julia} programming language, and argue this is necessary to
-obtain the three primary benefits of standardization mentioned earlier.
+These tools must be performant, but crucially modular and interfaceable
+(McIntire \emph{et al.} 2022)
 
@@
 these lower level langauges is not often taught.
+
+Ecological data is often difficult to access and reuse (Gonzalez \&
+Peres-Neto 2015; Poisot \emph{et al.} 2019). Synthesizing data into a
+single product suitable for analysis often remains tedious as data are
+not in formats that can be easily combined or interfaced.
+
+Here we propose that we can solve this problem through standardization
+(Zimmerman 2008)---developing a common definition such that data
+collected in a variety of contexts can be assimilated while minimizing
+the overhead of data cleaning and wrangling. A common representation of
+ecological data will have three primary benefits: it will \textbf{1})
+enable new forms of analysis by making it easier to combine data from
+different sources (Heberling \emph{et al.} 2021), \textbf{2)} enable
+continuous integration of new data for next-generation biodiversity
+monitoring (Kühl \emph{et al.} 2020), and \textbf{3)} aid in open
+sharing and reproducability of published results (Zimmerman 2008;
+Borregaard \& Hart 2016). Here, we briefly review approaches to data
+standardization developed in other fields, in order to determine what
+makes an open standard succeed in promoting data sharing, and what
+doesn't. Based on the properties of good standards we identify, we
+propose building a living standard for ecological data in the
+\texttt{Julia} programming language, and argue this is necessary to
+obtain the three primary benefits of standardization mentioned earlier.
 

### 2022-09-10 preprint.tex
Deploying to gh-pages from @ gottacatchenall/ms_julia_for_ecologists@738048d03f64b40771406b32f43895c9d9a41177 🚀
Deploying to gh-pages from @ gottacatchenall/ms_julia_for_ecologists@d199f6d27a498bbadc447c73c51cd36074336e9e 🚀

@@
 
-\noindent {\LARGE \textbf{\textsf{Solving the \(n\)-language problem in
-ecology}}}
+\noindent {\LARGE \textbf{\textsf{Solving the \(n\)-language problem: A
+ecologist's guide to learning Julia}}}
 
@@
 
+Ecological data is often difficult to access and reuse (Gonzalez \&
+Peres-Neto 2015; Poisot \emph{et al.} 2019). Synthesizing data into a
+single product suitable for analysis often remains tedious as data are
+not in formats that can be easily combined or interfaced.
+
+Here we propose that we can solve this problem through standardization
+(Zimmerman 2008)---developing a common definition such that data
+collected in a variety of contexts can be assimilated while minimizing
+the overhead of data cleaning and wrangling. A common representation of
+ecological data will have three primary benefits: it will \textbf{1})
+enable new forms of analysis by making it easier to combine data from
+different sources (Heberling \emph{et al.} 2021), \textbf{2)} enable
+continuous integration of new data for next-generation biodiversity
+monitoring (Kühl \emph{et al.} 2020), and \textbf{3)} aid in open
+sharing and reproducability of published results (Zimmerman 2008;
+Borregaard \& Hart 2016). Here, we briefly review approaches to data
+standardization developed in other fields, in order to determine what
+makes an open standard succeed in promoting data sharing, and what
+doesn't. Based on the properties of good standards we identify, we
+propose building a living standard for ecological data in the
+\texttt{Julia} programming language, and argue this is necessary to
+obtain the three primary benefits of standardization mentioned earlier.
+
 We propose that that Julia has certain properties absent in other
